SimTrans: A Self-supervised Method for Biomedical Image Segmentation

### 2023-06-19 cvpr_latex/PaperForReview.tex
add figure and table

To address the scarcity of biomedical image datasets, researchers aim to develop models that can learn features directly from the data without the need for labeled data. This approach, known as self-supervised learning, allows models to learn features from the data itself. One widely studied self-supervised learning framework in the field is SimCLR, proposed by Hinton et al.
SimCLR consists of four main components: data augmentation, a base encoder, a projection head, and a contrastive loss. The framework is designed to learn powerful representations from unlabeled data. In SimCLR, data augmentation is applied to generate multiple augmented versions of each unlabeled sample. These augmented samples are then passed through the base encoder, which extracts their features. The projection head further processes the encoded features to produce informative representations. Finally, the contrastive loss function encourages similar representations of augmented samples from the same sample while pushing representations from different samples apart.
The overall objective of SimCLR is to maximize agreement between differently augmented views of the same data sample while minimizing agreement between different samples. This enables the model to learn rich and discriminative representations without the need for explicit labels. SimCLR has garnered significant attention from researchers and has achieved remarkable results in various computer vision tasks, including biomedical image analysis.

@@

Swin Transformer is a hierarchical structure that builds feature maps by merging image patches at deeper layers, allowing flexible modeling at multiple scales. It achieves a linear computational complexity relative to the image size by computing self-attention within each local non-overlapping window. Additionally, Swin Transformer enables cross-window connections, as shown in Figure 2. The self-attention computation of a new window spans the boundaries of previously processed windows in layer l, providing cross-window connections between adjacent non-overlapping windows. This operation has been found effective in tasks such as image classification, object detection, and semantic segmentation. The ability to model visual entities of various scales and replace quadratic complexity with linear computational complexity contributes to Swin Transformer's versatility as a backbone for various visual tasks.
Swin Transformer consists of a Multi-head Self-Attention module (MSA) based on movable windows, followed by a two-layer Multilayer Perceptron (MLP) with Gaussian Error Linear Units (GELU) as the non-linear activation function. GELU is a high-performance neural network activation function that weights the inputs rather than gating them with signs as in ReLU. Layer Normalization (LN) is applied before each MSA module and each MLP, and a residual connection is employed after each module, as shown in Figure 3.
The computational complexity of the standalone MSA module in Swin Transformer has a quadratic relationship with the image size, as shown in Equation (3):
However, in Swin Transformer, the self-attention computation is performed within non-overlapping local windows, resulting in a linear computational complexity relative to the image size, as shown in Equation (4):
Here, M is a constant.
Furthermore, Swin Transformer incorporates relative position information into the attention computation, as shown in Equation (5):
Where Q, K, and V represent the query matrix, key matrix, and value matrix, respectively. d is the dimension, and B represents the relative position bias. Compared to using absolute positions, incorporating relative positional biases significantly improves the model's performance.
Additionally, Swin Transformer introduces inductive biases such as locality, hierarchical feature representation, and translation equivariance, enabling it to serve as a universal backbone network for various image recognition tasks.
Note: The translation assumes the correct representation of the formulas and symbols based on the original text.

@@

@@

@@

@@

@@

Figure 8 displays the following images from left to right: the original brain MRI image, feature maps from the first stage, feature maps from the second stage, feature maps from the third stage, feature maps from the fourth stage, and the ground truth label map. From the visualization, it can be observed that the SimTrans model progressively increases its attention to the lesion regions, demonstrating its effectiveness in segmenting the target regions.

The comparison results between SimTrans and two other networks, ResUNet and ViT, are presented in Table 1.

@@

@@

@@

\begin{table}
  \centering
  \begin{tabular}{@{}lc@{}}
    \toprule
    Method & Frobnability \\
    \midrule
    Theirs & Frumpy \\
    Yours & Frobbly \\
    Ours & Makes one's heart Frob\\
    \bottomrule
  \end{tabular}
  \caption{Results.   Ours is better.}
  \label{tab:example}
\end{table}

>>>>

To address the scarcity of biomedical image datasets, researchers aim to develop models that can learn features directly from the data without the need for labeled data. This approach, known as self-supervised learning, allows models to learn features from the data itself. One widely studied self-supervised learning framework in the field is SimCLR, proposed by Hinton et al.
SimCLR consists of four main components: data augmentation, a base encoder, a projection head, and a contrastive loss. The framework is designed to learn powerful representations from unlabeled data. In SimCLR, data augmentation is applied to generate multiple augmented versions of each unlabeled sample. These augmented samples are then passed through the base encoder, which extracts their features. The projection head further processes the encoded features to produce informative representations. Finally, the contrastive loss function encourages similar representations of augmented samples from the same sample while pushing representations from different samples apart.

\begin{figure*}
  \begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics{Images/SimCLR.png}
  \caption{Framwork of SimCLR.}
  \end{center}
  \label{fig:simclr}
\end{figure*}

The overall objective of SimCLR is to maximize agreement between differently augmented views of the same data sample while minimizing agreement between different samples. This enables the model to learn rich and discriminative representations without the need for explicit labels. SimCLR has garnered significant attention from researchers and has achieved remarkable results in various computer vision tasks, including biomedical image analysis.

@@

Swin Transformer is a hierarchical structure that builds feature maps by merging image patches at deeper layers, allowing flexible modeling at multiple scales. It achieves a linear computational complexity relative to the image size by computing self-attention within each local non-overlapping window. Additionally, Swin Transformer enables cross-window connections, as shown in Figure 2. The self-attention computation of a new window spans the boundaries of previously processed windows in layer l, providing cross-window connections between adjacent non-overlapping windows. This operation has been found effective in tasks such as image classification, object detection, and semantic segmentation. The ability to model visual entities of various scales and replace quadratic complexity with linear computational complexity contributes to Swin Transformer's versatility as a backbone for various visual tasks.

\begin{figure*}
  \begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics{Images/ST_1.png}
  \end{center}
     \caption{Self-attention computation for non-overlapping windows.}
  \label{fig:st_1}
\end{figure*}

Swin Transformer consists of a Multi-head Self-Attention module (MSA) based on movable windows, followed by a two-layer Multilayer Perceptron (MLP) with Gaussian Error Linear Units (GELU) as the non-linear activation function. GELU is a high-performance neural network activation function that weights the inputs rather than gating them with signs as in ReLU. Layer Normalization (LN) is applied before each MSA module and each MLP, and a residual connection is employed after each module, as shown in Figure 3.
The computational complexity of the standalone MSA module in Swin Transformer has a quadratic relationship with the image size, as shown in Equation (3):
However, in Swin Transformer, the self-attention computation is performed within non-overlapping local windows, resulting in a linear computational complexity relative to the image size, as shown in Equation (4):
Here, M is a constant.
Furthermore, Swin Transformer incorporates relative position information into the attention computation, as shown in Equation (5):
Where Q, K, and V represent the query matrix, key matrix, and value matrix, respectively. d is the dimension, and B represents the relative position bias. Compared to using absolute positions, incorporating relative positional biases significantly improves the model's performance.
Additionally, Swin Transformer introduces inductive biases such as locality, hierarchical feature representation, and translation equivariance, enabling it to serve as a universal backbone network for various image recognition tasks.

\begin{figure*}
  % \begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics[width=0.5\linewidth]{Images/ST_2.png}
  % \end{center}
  \caption{Block of Swin Transformer.}
  \label{fig:st_2}
\end{figure*}

@@

% \begin{figure*}
%   \begin{center}
%   % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}  
%   \includegraphics{Images/RCBA.png}
%   \end{center}
%      \caption{Structure of RCBA.}
%   \label{fig:rcba}
% \end{figure*}

@@

% \begin{figure*}
%   \begin{center}
%   % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
%   \includegraphics{Images/MSD.png}
%   \end{center}
%      \caption{Structure of MSD.}
%   \label{fig:msd}
% \end{figure*}

@@

\begin{figure*}
  \begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{Images/SimTrans.png}
  \end{center}
     \caption{Structure of SimTrans.}
  \label{fig:simtrans}
\end{figure*}

@@

\begin{figure*}
  \begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics{Images/DR_SimCLR.png}
  \end{center}
     \caption{Different representations of the same image produced by SimCLR.}
  \label{fig:dr_simclr}
\end{figure*}

@@

\begin{figure*}
  \begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics{Images/Vision.png}
  \end{center}
     \caption{Feature visualization.}
  \label{fig:vision}
\end{figure*}

Figure 8 displays the following images from left to right: the original brain MRI image, feature maps from the first stage, feature maps from the second stage, feature maps from the third stage, feature maps from the fourth stage, and the ground truth label map. From the visualization, it can be observed that the SimTrans model progressively increases its attention to the lesion regions, demonstrating its effectiveness in segmenting the target regions.
The comparison results between SimTrans and two other networks, ResUNet and ViT, are presented in Table 1.

\begin{table}
  \centering
  \begin{tabular}{lccc}
    \toprule
    Model & Dice & MIoU & Precision \\
    \midrule
    ResUNet & 0.6810 & 0.5552 & 0.5583 \\
    ViT & 0.6783 & 0.5496 & 0.5529 \\
    SimTrans & 0.7585 & 0.6424 & 0.6453 \\
    \bottomrule
  \end{tabular}
  \caption{Results.   Ours is better.}
  \label{tab:result_1}
\end{table}

@@

\begin{figure*}
  \begin{center}
  % \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \includegraphics{Images/Results.png}
  \end{center}
     \caption{Model segmentation visualization.}
  \label{fig:result}
\end{figure*}

@@

\begin{table}
  \centering
  \begin{tabular}{lccc}
    \toprule
    Metric & ResUNet & ViT & SimTrans \\
    \midrule
    Parameters/MB & 24.01 & 108.25 & 87.32 \\
    Complexity/GMac & 15.24 & 55.42 & 47.51 \\
    \bottomrule
  \end{tabular}
  \caption{Results.   Ours is better.}
  \label{tab:result_2}
\end{table}

@@
